---
title: "Regression Modeling"
author: "Group Healthcare Crisis: Shrey, Cameron, Yoel"
date: "`r Sys.Date()`"
output: 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning=FALSE, message=FALSE}
# Package prep
require(tidyr)
require(dplyr)
require(tibble)
require(readr)
require(mosaic)
require(ggplot2)
require(reshape2)
require(leaps)
require(lubridate)
library(corrplot)
library(leaps)
library(MASS)
library(openxlsx)

# Data prep
<<<<<<< Updated upstream
data <- read_csv("0511_merged.csv")
=======
data <- read_csv("https://www.dropbox.com/s/72datd3ntg7ey2z/final_merged_data.csv?dl=1")
>>>>>>> Stashed changes
```
## Data

```{r}
reduced <-data
reduced <- dplyr::select(data, -c(fips, week, cases, deaths, black, hispanic, native, asian, white))
```

## Correlation Plot

```{r}
plot.new()
dev.off()

# Select all columns except "CFR"
data_clean_without_CFR <- na.omit(reduced[, !colnames(reduced) %in% "CFR"])

# Compute the correlation matrix
mycorr <- cor(data_clean_without_CFR)

# Plot the correlation matrix
corrplot(mycorr, type = "upper", order = "hclust", tl.col = "black", tl.cex = 0.45, method="shade")



```

<<<<<<< Updated upstream
=======
## Pairs Plot

```{r}
#ncol(data)

# Set the plot margins to a smaller value
#par(mar = c(0.002, 0.002, 0.002, 0.002))

# Create the pairs plot
#pairs(data[,1:36], main = "Pairs plot of 36 columns of data")

# Create a pairs plot
#pairs(data, main = "Pairs Plot for Covid 2021 Data Set")
```




>>>>>>> Stashed changes
## Best subsets
```{r}
## Dividing the large dataset
half1 <- reduced[, -(19:36)]
half2 <- reduced[, -(2:18)]

# Load the "leaps" package
library(leaps)

# Perform best subsets regression on dataset 1
best_subsets1 <- regsubsets(CFR ~ ., data = half1, nvmax = ncol(half1)-1, method = "exhaustive")
summary(best_subsets1, rsq = TRUE)

# Perform best subsets regression on dataset 2
best_subsets2 <- regsubsets(CFR ~ ., data = half2, nvmax = ncol(half2)-1, method = "exhaustive")
summary(best_subsets2, rsq = TRUE)
```
## Best Subsets - 1st half plot

```{r}
<<<<<<< Updated upstream
par(fig=c(0, 1, 0, 1), mar=c(9, 4, 6, 9) + 0.1, oma=c(0, 0, 2, 0), cex.axis=1.18, cex.lab=0.8)
par(fig=c(0, 1, 0.35, 1), new=TRUE)

# Modify variable names
modified_names <- gsub("_7_day_avg", "", best_subsets1$xnames)

# Plot the best subsets without default x-axis labels
plot(best_subsets1, scale="r2", xaxt="n", labels = FALSE)

# Add custom x-axis labels
axis(1, at=1:length(modified_names), labels=modified_names, las=2)
=======
# Set plot size and margins


# Plot the best subsets
plot(best_subsets1, scale="r2", las=2)

# Add a title and adjust the outer margin
mtext("1st half", outer=TRUE, line=0.5, cex=1.2)
>>>>>>> Stashed changes

"pop_over_65"
"under18"
"commute"
"health_insur_coverage"
"household_income"
"male"
"income_below_poverty"
"total_staffed_adult_icu_beds_7_day_avg"

# Set plot size and margins


```
<<<<<<< Updated upstream

## Best Subsets - 2nd half plot

```{r}
par(fig=c(0, 1, 0, 1), mar=c(9, 4, 6, 9) + 0.1, oma=c(0, 0, 2, 0), cex.axis=1.18, cex.lab=0.8)
par(fig=c(0, 1, 0.35, 1), new=TRUE)

# Modify variable names
modified_names <- gsub("_7_day_avg", "", best_subsets2$xnames)

# Plot the best subsets without default x-axis labels
plot(best_subsets2, scale="r2", xaxt="n", labels = FALSE)

# Add custom x-axis labels
axis(1, at=1:length(modified_names), labels=modified_names, las=2)

=======
## Best Subsets - 2nd half plot

```{r}
# Set plot size and margins


# Plot the best subsets
plot(best_subsets2, scale="r2", las=2)

# Add a title and adjust the outer margin
mtext("2nd half", outer=TRUE, line=0.5, cex=1.2)
>>>>>>> Stashed changes
"staffed_icu_adult_patients_confirmed_covid_7_day_avg"
"binge"
"casthma"
"chd"
"csmoking"
"diabetes"
"lpa"
"obesity"
"Administered_Dose1_Pop_Pct"
```
## Best Subsets - final

```{r}
finalhalf <- dplyr::select(reduced, CFR, staffed_icu_adult_patients_confirmed_covid_7_day_avg, binge, casthma, chd, csmoking, diabetes, lpa, obesity, Administered_Dose1_Pop_Pct, pop_over_65, under18, commute, health_insur_coverage, household_income, male, income_below_poverty, total_staffed_adult_icu_beds_7_day_avg)
    
best_subsets3 <- regsubsets(CFR ~ ., data = finalhalf, nvmax = ncol(half1)-1, method = "exhaustive")
summary(best_subsets1, rsq = TRUE)

<<<<<<< Updated upstream
par(fig=c(0, 1, 0, 1), mar=c(9, 4, 6, 9) + 0.1, oma=c(0, 0, 2, 0), cex.axis=1.18, cex.lab=0.8)
par(fig=c(0, 1, 0.35, 1), new=TRUE)

# Modify variable names
modified_names <- gsub("_7_day_avg", "", best_subsets3$xnames)

# Plot the best subsets without default x-axis labels
plot(best_subsets3, scale="r2", xaxt="n", labels = FALSE)

# Add custom x-axis labels
axis(1, at=1:length(modified_names), labels=modified_names, las=2)
=======
# Set plot size and margins
par(fig=c(0, 1, 0, 1), mar=c(5, 4, 8, 2) + 0.1, oma=c(0, 0, 2, 0), cex.axis=0.8, cex.lab=0.8)
par(fig=c(0, 1, 0.25, 1), new=TRUE)

# Plot the best subsets
plot(best_subsets3, scale="r2", las=2)

# Add a title and adjust the outer margin
mtext("Final Best Subsets", outer=TRUE, line=0.5, cex=1.2)
>>>>>>> Stashed changes
```

## Final Subset Model
```{r}
finalhalf <- dplyr::select(reduced, CFR, staffed_icu_adult_patients_confirmed_covid_7_day_avg, binge, casthma, chd, csmoking, diabetes, lpa, obesity, Administered_Dose1_Pop_Pct, pop_over_65, under18, commute, health_insur_coverage, household_income, male, income_below_poverty, total_staffed_adult_icu_beds_7_day_avg)
    
#best_subsets3 <- regsubsets(CFR ~ ., data = finalhalf, nvmax = ncol(half1)-1, method = "exhaustive")
#summary(best_subsets1, rsq = TRUE)
#plot(best_subsets3, scale="adjr2")

regfin <- dplyr::select(reduced, CFR, staffed_icu_adult_patients_confirmed_covid_7_day_avg, binge, chd, csmoking, diabetes, lpa, pop_over_65, under18, commute, health_insur_coverage, male)

reg1 <- lm(CFR ~., data = regfin)
summary(reg1)
```

#resid plots
```{r}
dev.new(width=12, height=16)
par(mfrow=c(5, 4), mar=c(2, 2, 2, 2))
for(i in 2:length(names(reg1$model))) {
  plot(reg1$model[,i], resid(reg1), main=paste("Residuals vs.", names(reg1$model)[i]), ylab="Residuals")
    # Add a horizontal line at y=0
  abline(h = 0, col = "red")
}
```
## npp
```{r}
dev.new(width=12, height=16)
par(mfrow=c(5, 4), mar=c(2, 2, 2, 2))
for(i in 2:length(names(reg1$model))) {
  qqnorm(reg1$model[,i], main=paste("Normal Probability Plot of", names(reg1$model)[i]))
  qqline(reg1$model[,i])
}
```

# DOING LOG TRANSFORMATION ON VARIABLES WITH FUNNELL SHAPED RESID PLOTS
```{r}
# Load the necessary libraries
library(dplyr)

# Remove NAs and zeroes from commute and staffed_icu_adult_patients_confirmed_covid_7_day_avg
regfin2 <- regfin %>%
  filter(commute > 0 & !is.na(commute)) %>%
  filter(staffed_icu_adult_patients_confirmed_covid_7_day_avg > 0 & !is.na(staffed_icu_adult_patients_confirmed_covid_7_day_avg)) %>%
  filter(health_insur_coverage > 0 & !is.na(health_insur_coverage)) %>%
  filter(male > 0 & !is.na(male)) %>%
  filter(pop_over_65 > 0 & !is.na(pop_over_65))

# Take the log of commute and staffed_icu_adult_patients_confirmed_covid_7_day_avg
regfin2$log_commute <- log(regfin2$commute)
regfin2$log_staffed_icu <- log(regfin2$staffed_icu_adult_patients_confirmed_covid_7_day_avg)
regfin2$log_pop_over_65 <- log(regfin2$pop_over_65)
regfin2$log_male <- log(regfin2$male)
regfin2$log_health_insur_coverage <- log(regfin2$health_insur_coverage)

# Create the new linear regression model with the logged variables
<<<<<<< Updated upstream
reg2 <- lm(CFR ~ binge + csmoking + diabetes + lpa + log_male + log_pop_over_65 + under18 + log_health_insur_coverage + log_commute + log_staffed_icu + csmoking*log_commute + diabetes*under18 + lpa*log_pop_over_65 + under18*log_commute + log_health_insur_coverage*log_commute + csmoking*binge + binge*diabetes, data = regfin2)
=======
reg2 <- lm(CFR ~ binge + chd + csmoking + diabetes + lpa + log_male + log_pop_over_65 + under18 + log_health_insur_coverage + log_commute + log_staffed_icu + csmoking*log_commute + diabetes*under18 + lpa*log_pop_over_65 + under18*log_commute + log_health_insur_coverage*log_commute + csmoking*binge + binge*diabetes + chd*diabetes + chd*lpa, data = regfin2)
>>>>>>> Stashed changes

# Print the summary of the new regression model
summary(reg2)
```

```{r}

# Full Model

<<<<<<< Updated upstream
regfin3 <- dplyr::select(data, CFR, staffed_icu_adult_patients_confirmed_covid_7_day_avg, binge, chd, csmoking, diabetes, lpa, pop_over_65, under18, commute, health_insur_coverage, male, white)
=======
regfin3 <- dplyr::select(data, CFR, staffed_icu_adult_patients_confirmed_covid_7_day_avg, binge, chd, csmoking, diabetes, lpa, pop_over_65, under18, commute, health_insur_coverage, male, white, black, hispanic, native, asian)
>>>>>>> Stashed changes

regfin3 <- regfin3 %>%
  filter(commute > 0 & !is.na(commute)) %>%
  filter(staffed_icu_adult_patients_confirmed_covid_7_day_avg > 0 & !is.na(staffed_icu_adult_patients_confirmed_covid_7_day_avg)) %>%
  filter(health_insur_coverage > 0 & !is.na(health_insur_coverage)) %>%
  filter(male > 0 & !is.na(male)) %>%
  filter(pop_over_65 > 0 & !is.na(pop_over_65))

regfin3$log_commute <- log(regfin3$commute)
regfin3$log_staffed_icu <- log(regfin3$staffed_icu_adult_patients_confirmed_covid_7_day_avg)
regfin3$log_pop_over_65 <- log(regfin3$pop_over_65)
regfin3$log_male <- log(regfin3$male)
regfin3$log_health_insur_coverage <- log(regfin3$health_insur_coverage)

<<<<<<< Updated upstream
reg3 <- lm(CFR ~ white + binge + csmoking + diabetes + lpa + log_male + log_pop_over_65 + under18 + log_health_insur_coverage + log_commute + log_staffed_icu + csmoking*log_commute + diabetes*under18 + lpa*log_pop_over_65 + under18*log_commute + log_health_insur_coverage*log_commute + csmoking*binge + binge*diabetes, data = regfin3)

summary(reg3)

# Extra Sum of Squares Test

=======
reg3 <- lm(CFR ~ white + black + hispanic + native + asian + binge + chd + csmoking + diabetes + lpa + log_male + log_pop_over_65 + under18 + log_health_insur_coverage + log_commute + log_staffed_icu + csmoking*log_commute + diabetes*under18 + lpa*log_pop_over_65 + under18*log_commute + log_health_insur_coverage*log_commute + csmoking*binge + binge*diabetes + chd*diabetes + chd*lpa, data = regfin3)

summary(reg3)
regre3 <- summary(reg3)
cc <- regre3$coefficients
term <- regre3$aliased

library(MASS)
eqn <- paste("Y =", paste(round(cc[1],2), paste(round(cc[-1],2), names(term[-1]), sep=" * ", collapse=" + "), sep=" + "), "+ e")
eqn

# Extra Sum of Squares Test
# black + hispanic + native + asian + white*chd + white*csmoking + b + 
>>>>>>> Stashed changes

# Calculate the extra sum of squares
anova(reg2, reg3)
```


# RESID AFTER LOG
```{r}
dev.new(width=12, height=16)
par(mfrow=c(5, 4), mar=c(2, 2, 2, 2))
for(i in 2:length(names(reg2$model))) {
  plot(reg2$model[,i], resid(reg2), main=paste("Residuals vs.", names(reg2$model)[i]), ylab="Residuals")
    # Add a horizontal line at y=0
  abline(h = 0, col = "red")
}
```
#Npp PLOTS AFTER
```{r}
dev.new(width=12, height=16)
par(mfrow=c(5, 4), mar=c(2, 2, 2, 2))
for(i in 2:length(names(reg2$model))) {
  qqnorm(reg2$model[,i], main=paste("Normal Probability Plot of", names(reg2$model)[i]))
  qqline(reg2$model[,i])
}

```

## VIF
```{r}
# Load the car and ggplot2 packages
library(car)
library(ggplot2)

# Fit a linear regression model using all predictor variables in the data frame
<<<<<<< Updated upstream
=======
regfin <- regfin %>% 
  rename(staffed_icu_covid = 
           staffed_icu_adult_patients_confirmed_covid_7_day_avg)
>>>>>>> Stashed changes
model <- lm(CFR ~ ., data = regfin)
summary(model)

# Calculate the VIF for each predictor variable
vif_values <- vif(model)

# Create a data frame with the predictor variable names and VIF values
df_vif <- data.frame(predictor = names(vif_values), vif = vif_values)

# Create a bar plot of the VIF values
<<<<<<< Updated upstream
=======
tiff("/Users/cameronlian/Downloads/Rplot.tiff", units="in", width=5, height=2, res=300)
>>>>>>> Stashed changes
ggplot(data = df_vif, 
       mapping = aes(x = predictor, 
                     y = vif)) +
  geom_col() +
<<<<<<< Updated upstream
  geom_hline(yintercept = 5, lty = 2, lwd = 2) +
  scale_x_discrete(guide = guide_axis(check.overlap = TRUE, 
                                      n.dodge = 3)) +
  ylab("Variance inflation factor (VIF)") +
  theme(axis.text = element_text(size = rel(1.7)), 
        axis.title = element_text(size = rel(2)), 
        axis.title.x = element_text(margin = margin(t = 10)), 
        axis.title.y = element_text(vjust = 1.5))
=======
  geom_hline(yintercept = 5, lty = 2, lwd = 0.8) +
  scale_x_discrete(guide = guide_axis(check.overlap = TRUE, 
                                      n.dodge = 3)) +
  ylab("Variance inflation factor (VIF)") +
  theme_classic() +
  theme(axis.text = element_text(size = rel(1)), 
        axis.title = element_text(size = rel(1)), 
        axis.title.x = element_text(size = rel(0.6)), 
        axis.text.x = element_text(size = rel(0.7)), 
        axis.title.y = element_text(size = rel(0.6)))
dev.off()
>>>>>>> Stashed changes

regfin <- dplyr::select(reduced, CFR, staffed_icu_adult_patients_confirmed_covid_7_day_avg, binge, chd, csmoking, diabetes, lpa, pop_over_65, under18, commute, health_insur_coverage, male)

lm_vif <- lm(CFR ~., data = regfin)
summary(lm_vif)

```

